---
title: "Valiant - Analysis financed emissions of mortgages"
author: "Annika Schmidt"
date: "2024-10-23"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "https://cloud.r-project.org/"))
```

# LOAD PACKAGES 
```{r, message = FALSE}
# LOAD PACKAGES ############################################
install.packages("pacman")
install.packages("openxlsx")
install.packages("knitr")
pacman::p_load(pacman, dplyr, GGally, ggplot2, ggthemes,
               ggvis, httr, lubridate, plotly, rio, rmarkdown, shiny,
               stringr, tidyr)
library(datasets)
library(openxlsx)
library(ggplot2)
library(dplyr)
library(knitr)
```
# IMPORT FILE 
```{r, message = FALSE}
# Excel XLSX
INTERACTIVE <- TRUE
if (INTERACTIVE) {
  file_in <- file.choose() # Use File Explorer
} else {
  file_in <- excel_filepath
}
wb <- openxlsx2::wb_load(file_in)

# Read the excel file
data <- openxlsx2::wb_to_df(wb, sheet = "byBuilding", start_row = 5, col_names = TRUE)
data_subset <- data[1:23380, ]
# Check the new dimensions
cat("Number of rows in the subset:", nrow(data_subset), "\n")
```

# PRE-ANALYSIS
## ERROR COMMENTS

```{r}
# Group by ErrorComments and count occurrences
error_summary <- data_subset %>%
  filter(!is.na(error_comments)) %>%
  group_by(error_comments) %>%
  summarise(Count = n()) %>%
  arrange(desc(Count))

# View the summary
kable(error_summary, caption = "Summary of Error Comments")
```

## cHECK FOR OUTLIERS
```{r, echo=FALSE}
# Boxplots
#It displays the distribution of data based on quartiles and highlights outliers as points outside the whiskers.
ggplot(data_subset, aes(y = emissions_per_area)) +
  geom_boxplot(fill = "lightblue", color = "black") +
  labs(title = "Boxplot with Outliers", y = "Emmissions per area") +
  theme_minimal()
```

```{r, echo=FALSE}
# Density plot
ggplot(data_subset, aes(x = emissions_per_area)) +
  geom_density(fill = "lightblue", alpha = 0.5) +
  geom_rug(sides = "b", color = "red") +
  labs(title = "Density Plot with Outliers", x = "Emissions per area") +
  theme_minimal()
```
### FILTER OUTLIERS
```{r}
# Mutate observations where 'emissions_per_area' is greater 150
data_subset_reduced <- data %>%
  mutate(emissions_per_area = ifelse(emissions_per_area > 150, NA, emissions_per_area))
```

## CREATE VARIABLES
```{r}
# Create a subset where mortgage_value is not NA
data_subset_non_na_mortgage <- data_subset %>%
  filter(!is.na(mortgage_emissions))

# Total energetic surface of the assets (m2)
total_asset_energetic_area <- sum(data_subset$asset_energetic_area, na.rm = TRUE)

# total asset value (CHF)
total_asset_value <- sum(data_subset$asset_value, na.rm = TRUE)

# total mortgage value (CHF)
total_mortgage_value <- sum(data_subset$mortgage_value, na.rm = TRUE)
total_mortgage_value_subset <- sum(data_subset_non_na_mortgage$mortgage_value, na.rm = TRUE)

# total asset emissions (absolute, in kg CO2)
total_asset_emissions <- sum(data_subset$asset_emissions, na.rm = TRUE)
# total financed emissions (absolute, in kg CO2)
# variable already multiplied by bank share
total_financed_emissions <- sum(data_subset$mortgage_emissions, na.rm =TRUE)

# Mean emission intensity (t CO2eq / Mio. CHF invested)
mean_emission_intensity_per_investment <- (sum(data_subset$mortgage_emissions, na.rm = TRUE)/1000)/(total_mortgage_value_subset/1000000)

# Mean emission intensity (kg CO2/ m2)
mean_emission_intensity_per_m2 <- mean(data_subset$emissions_per_area, na.rm = TRUE)

# mean bank share of mortgage on total assets
mean_bank_share <- total_mortgage_value / total_asset_value
```

```{r}
options(scipen = 999)
# Create a data frame to summarize all the variables
summary_variables <- data.frame(
  Metric = c("Total Asset Energetic Area (m2)",
             "Total Asset Value (CHF)",
             "Total Mortgage Value (CHF)",
             "Total Asset Emissions (kg CO2)",
             "Total Financed Emissions (kg CO2)",
             "Mean Emission Intensity per Mio CHF (t CO2eq/ Mio. CHF invested)",
             "Mean Emission Intensity per m2 (kg CO2/m2)",
             "Mean Bank Share of Mortgage on Total Assets"),
  Value = c(total_asset_energetic_area,
            total_asset_value,
            total_mortgage_value,
            total_asset_emissions,
            total_financed_emissions,
            mean_emission_intensity_per_investment,
            mean_emission_intensity_per_m2,
            mean_bank_share)
)
# Print the data frame
print(summary_variables)

```

# ANALYSIS

## ANALYSIS BY 
KPI SUMMARY grouped by different variables

```{r}
gklas_mapping <- data.frame(
  GKLAS = c(1110, 1121, 1122, 1130, 1211, 1212, 1220, 1230, 1242, 1251),
  GKLAS_def = c('Maisons à un logement', 'Maisons à deux logements', 'Immeubles à trois logements et plus', 'Habitat communautaire', 'Hôtels', 'Autres bâtiments dhébergement de tourisme', 'Immeubles de bureaux', 'Bâtiments commerciaux', 'Garages', 'Bâtiments industriels')  # Replace with your specific numbers
)

# Join the mapping table with the original data frame
data_subset <- data_subset %>%
  left_join(gklas_mapping, by = "GKLAS")

```

### 3) GKLAS
GKLAS = Classe de bâtiment selon EUROSTAT (voir RegBL_Catalogue, page 57/58)
```{r}
type_summary_GKLAS <- data_subset %>%
  group_by(GKLAS_def) %>%
  summarise(
    mean_emissions = mean(emissions_per_area, na.rm = TRUE),
    mean_emission_intensity_per_investment = sum(mortgage_emissions, na.rm = TRUE)/1000/(sum(mortgage_value[!is.na(mortgage_emissions)], na.rm = TRUE)/1000000),
    sum_financed_emissions = sum(asset_emissions, na.rm = TRUE),
    sum_asset_energetic_area = sum(asset_energetic_area, na.rm = TRUE),
    share_mortgage_value = (sum(mortgage_value, na.rm = TRUE)/ total_mortgage_value),
    count = n()
  )
print(type_summary_GKLAS, caption = "Emission Summary by GKLAS")
```
### 4) ENERGY CARRIER
```{r}
type_summary_energy_carrier <- data_subset %>%
  group_by(sia_energy_carrier) %>%
  summarise(
    mean_emissions = mean(emissions_per_area, na.rm = TRUE),
    mean_emission_intensity_per_investment = sum(mortgage_emissions, na.rm = TRUE)/1000/(sum(mortgage_value[!is.na(mortgage_emissions)], na.rm = TRUE)/1000000),
    sum_financed_emissions = sum(asset_emissions, na.rm = TRUE),
    sum_asset_energetic_area = sum(asset_energetic_area, na.rm = TRUE),
    share_mortgage_value = (sum(mortgage_value, na.rm = TRUE)/ total_mortgage_value),
    count = n()
  )
print(type_summary_energy_carrier, caption = "Emission Summary by Energy Carrier")
```

Now, export all summaries to Excel
```{r}
wb2 <- createWorkbook()
# List of data frames (tables) to export
tables_list <- list(
  "GKLAS Summary" = type_summary_GKLAS,
  "Energy Carrier Summary" = type_summary_energy_carrier
)

# Loop over the list of tables and export each as a sheet
for (sheet_name in names(tables_list)) {
  addWorksheet(wb2, sheet_name) # Create a sheet with the table name
  writeData(wb2, sheet = sheet_name, tables_list[[sheet_name]]) # Write the table to the corresponding sheet
}

# Save the workbook to an Excel file
saveWorkbook(wb2, "Emission_Summary_Tables_BCVs.xlsx", overwrite = TRUE)
```

# COVERAGES
## 1) Relative to number of mortgages
Comparing the number of non-missing values for each variable to the total number of rows (one row = one mortgage)
```{r}
# Addresses available = proportion of addresses for which we have valid data.
total_rows <- nrow(data_subset)
print(total_rows)

# Coverage ratio for EGID (non-missing EGID)
coverage_EGID <- sum(!is.na(data_subset$EGID)) / total_rows
print(coverage_EGID)

# Coverage ratio for emisisons_per_area
coverage_mortgage_emissions <- sum(!is.na(data_subset$mortgage_emissions)) / total_rows
print(coverage_mortgage_emissions)
```

## 2) Coverage ratio for financed emissions relative to investments (%)
```{r}
# where do I have data for mortage emissions, * mortgage_value / total_mortgage_value
data_subset <- data_subset %>%
  mutate(
    emissions_proportion = ifelse(is.na(mortgage_emissions),
                                  mortgage_value,
                                  NA)
  )
coverage_financed_emissions <- sum(data_subset$emissions_proportion, na.rm = TRUE) / total_mortgage_value
print(1-coverage_financed_emissions)
```
Whereas the code can get results for 71% of the rows, we have results for 72% of total mortgage value.

## 3) Coverage ratio related to area
Sum of energy relevant area when financed emissions is not NA / Total energy relevant area
```{r}
coverage_ratio_area <- data_subset %>%
  summarise(
    sum_area_condition = sum(asset_energetic_area[!is.na(mortgage_emissions)], na.rm = TRUE),
    sum_area = sum(asset_energetic_area, na.rm = TRUE)
  ) %>%
  mutate(ratio = sum_area_condition / sum_area)

print(coverage_ratio_area)
```

# PLOTS
According to years built
```{r, message=FALSE}
ggplot(data = data_subset, aes(x = sia_year)) +
  geom_histogram(binwidth = 10, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Building Construction Years",
       x = "Year of Construction",
       y = "Number of Buildings") +
  scale_x_continuous(breaks = seq(1900, 2030, by = 10), limits = c(1900,2030)) +
  scale_y_continuous(breaks = seq(0, 35000, by = 1000)) + 
  theme_minimal() # Adjust breaks as needed
```

Distribution of emissions per area (kg CO2/m2)
```{r, message=FALSE}
ggplot(data = data_subset, aes(x = emissions_per_area)) +
  geom_histogram(binwidth = 10, fill = "red", color = "black") +
  labs(title = "Emission Intensity per Square Meter for Mortgaged Properties",
       x = "Emission Intensity (kg CO2/m2)",
       y = "Number of Properties") +
  scale_x_continuous(breaks = seq(0, 400, by = 10)) +
  scale_y_continuous(breaks = seq(0, 35000, by = 1000)) + # Adjust breaks as needed
  theme_minimal()
```

Share of different energy carriers
```{r, message=FALSE}
data_subset$energy_carrier[data_subset$sia_energy_carrier == "4"] <- NA ##
energy_carrier_counts <- table(data_subset$sia_energy_carrier)
custom_names <- c("Gas","Oil","Other")

pie(energy_carrier_counts,
    labels = paste(custom_names, round(prop.table(energy_carrier_counts) * 100, 1), "%"),
    col = c("blue", "green", "red"),
    main = "Proportional Distribution of Energy Carriers")

```
