---
title: "BuildingsEmissionsCalculator"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BuildingsEmissionsCalculator}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Getting started

## Install R
[Download and install R](https://cran.rstudio.com/)

## Install RStudio IDE (optional)
This step is optional. An other IDE or none at all can be used.  
[Download and install RStudio](https://posit.co/download/rstudio-desktop/)

## Install packages

### From R command line

```{r, eval=FALSE}
install.packages("C:/path/to/co2calculatorPACTA2022_1.2.2.tar.gz", repos = NULL, type = "source")
install.packages("C:/path/to/BuildingsEmissionsCalculator_0.0.1.tar.gz", repos = NULL, type = "source")
```

### From RStudio

- Click `Tools → Install Packages`
- Select `Package Archive File (.zip, .tar.gz)` in the `Install from:` slot
- Find the corresponding file on the local machine, and click `Open`
- Click `Install`

### Run BuildingsEmissionsCalculator

Run the following R code 

```{r, eval=FALSE}
library(co2calculatorPACTA2022)
library(BuildingsEmissionsCalculator)

# Choose if the File Explorer is used to seach Excel file
INTERACTIVE <- FALSE

# Load the Excel file
if (INTERACTIVE) {
  file_in <- file.choose() # Use File Explorer
} else {
  file_in <- "path/to/building_data.xlsx"
}
buildings_df <- readxl::read_excel(file_in, n_max = 10)

# Call the BuildingsEmissionsCalculator
buildings_df <- fill_buildings_df(buildings_df)

# Save the data to a Excel file
if (INTERACTIVE) {
  file_out <- file.choose() # Use File Explorer
} else {
  file_out <- sub("\\.xlsx$", "_filled.xlsx", file_in)
}
writexl::write_xlsx(buildings_df, file_out)
```


